---
title: "covid 19"
author: "Kaylie Brehm"
date: "Summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```

  
# Introduction 

Question 1 - Is dying independent of sex?

Investigate whether men and women are equally likely to die of covid19 from data collected in south korea through May 2019.

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May.

Some of the data was uncoded for sex - sex must be just male or female.
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well.


# Method


$H_0$: In South Korea through May 2019, death from covid19 is not affected from sex.

$H_a$: In South Korea through May 2019, death from covid19 is affected from sex.

Mortality (alive or dead) is a categorical variable. Sex (male or female) is also a categorical variable. We will use a CAT ~ CAT set of graphs to determine whether death from covid19 was impacted by sex in South Korea through May 2019. I will use a gg style bar chart, one showing frequency and one showing percent. I will also use a numerical findings chart showing percentages and observed counts. Finally I will use a chi-squared test and a fisher exact test.

# Results

## Graphical Results

We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

````{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


The data shows that more women are alive than men. However, the percentage of females who are alive is more than the percentage of males that are alive. As for the dead, it shows that the percentage of males that are dead is about twice as higher as the percentage of females that are dead. 


## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```


Out of the total population of people who are alive, 56% of them are females while 44% of them are males. Out of the total population of people who are dead, 36% of them are females while 64% of them are males.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The chi sq test provides the exact numbers. It shows that there were 1831 alive females and 1411 alive males. 25 women died vs 44 males died. The odds of a female dying is (25/1831). The odds of a male dying is (44/1411). The odds ratio of a male dying to a female is (44/1411)/(25/1831). The odds ratio is 2.3 to 1, meaning men are 2.3 times more likely to die. We also see the confidence intervals are, on the low end 1.4 and 3.9 on the high end. The chi-square test shows a p-value of 0.0012. The fisher test shows a p-value of 0.0000000004. Since these are both under 0.05, we reject the null hypothesis.


# Conclusions


We investigated whether men and women are equally likely to die of covid19 from data collected in
South Korea through May 2019. The null hypothesis was "In South Korea through May 2019, death from covid19 is not affected from sex." The numerical summary showed that out of the population of people who were dead, 64% of them were males while only 36% of them were females. When looking at the Fisher's Exact Test, we can see that the odds ratio of dying is 2.3 to 1 if you are a male. Finally, since the p-value is under 0.05, at 0.0000000004, we reject the null hypothesis. I can conclude that there is a correlation between sex and dying. 

