---
title: "covid 19"
author: "Kaylie Brehm"
date: "Summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```

### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Introduction 

Question 1 - Is dying independent of sex?


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

Investigate whether men and women are equally likely to die of covid19 from data collected in south korea through May 2019



# Method


$H_0$: In South Korea through May 2019, death from covid19 is not affected from sex.

$H_a$: In South Korea through May 2019, death from covid19 is affected from sex.


#Results

## Graphical Results

We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

````{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


# Conclusions

